A Parallel, Backjumping Subgraph Isomorphism Algorithm using Supplemental Graphs

### 2015-06-20 paper/paper.tex
Better spacing

\begin{abstract}
    The subgraph isomorphism problem involves finding a pattern graph inside a target graph.  We
    present a new bit- and thread-parallel constraint-based search algorithm for the problem, and
    experiment on a wide range of standard benchmark instances to demonstrate its effectiveness.  We
    introduce the idea of supplemental graphs, to create implied constraints. We use a low-overhead,
    lazy variation of conflict directed backjumping which interacts safely with parallel search, and
    a counting-based all-different propagator which is better suited to our large domains than the
    conventional matching approach.
\end{abstract}

@@

This constraint-based search approach dates back to works by Ullmann \cite{Ullmann:1976} and
McGregor \cite{McGregor:1979}, and was improved upon in the LV \cite{Larrosa:2002} and VF2
\cite{Cordella:2004} algorithms.  More recently, ILF \cite{Zampelli:2010}, LAD \cite{Solnon:2010}
and SND \cite{Audemard:2014} are algorithms which take a ``deep thinking'' approach, using strong
inference at each stage of the search.  This is powerful, but we observe that when LAD or SND fail
to find a solution in time, it is often because they do not get to explore much of the search space.
Sometimes LAD and SND make less than one recursive call per second---this is often true with larger
target graphs, where strong inference becomes particularly expensive.  This motivates our
alternative approach: on the same hardware, we will be making between ten thousand and a million
recursive calls per core per second on non-trivial instances.  The main features of our algorithm
are:

@@

    \item We describe a clone-comparing variation of conflict-directed backjumping, which does not
        require explicit conflict sets. We note that an all-different propagator can produce reduced
        conflict sets, which can enable more backjumping.

@@

\begin{proposition}\label{proposition:supplemental}
    Let $i : P \hookrightarrow T$ be a subgraph isomorphism. Then $i$ is also
    \begin{enumerate}[itemindent=1em,topsep=0pt,itemsep=0.2pt]
        \item a subgraph isomorphism $i^d : P^d \hookrightarrow T^d$ for any $d \ge 1$, and
        \item a subgraph isomorphism $i^{\left[c, l\right]} : P^{\left[c, l\right]} \hookrightarrow
            T^{\left[c, l\right]}$ for any $c, l \ge 1$.
    \end{enumerate}
\end{proposition}

@@

\begin{algorithm}[h]
\DontPrintSemicolon
\nl $\FuncSty{nonInducedSubgraphIsomorphism}$ (Graph $\mathcal{P}$, Graph $\mathcal{T}$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$|\operatorname{V}(\mathcal{P})| > |\operatorname{V}(\mathcal{T})|$}{$\KwSty{return}~\KwSty{false}$\label{line:enough}}
    \nl Discard isolated vertices in $\mathcal{P}$\label{line:isolated} \;
    \nl $L \gets \FuncSty{createSupplementalGraphList}(\mathcal{P}, \mathcal{T})$ \label{line:buildsup}\;
    \nl $D \gets \FuncSty{init}(\operatorname{V}(\mathcal{P}), \operatorname{V}(\mathcal{T}), L)$
    \label{line:initdomains}\;
    \nl \lIf{$\FuncSty{countingAllDifferent}(D) \ne \textnormal{Success}$}{$\KwSty{return}~\KwSty{false}$\label{line:initalldiff}}
    \nl $\KwSty{return}~\FuncSty{search}(L, D) = \textnormal{Success}$ \label{line:search} \;
}

@@

\begin{algorithm}[h]
\DontPrintSemicolon
\nl $\FuncSty{init}$ (Vertices $V$, Vertices $R$, GraphPairs $L$) $\rightarrow$ Domains \;
\nl \Begin{
    \nl \Repeat{\nl$R$ \textnormal{is unchanged}}{
        \nl \ForEach{$v \in V$}{
            \nl $D_v \gets \bigcap_{(P,\,T) \in L} \left\{ w \in R :~
                v\underset{P}{\sim}v \Rightarrow w\underset{T}{\sim}w \wedge \operatorname{S}(P, v) \preceq
            \operatorname{S}(T[R], w) \right\}$ \label{line:filter} \;
        }
    \nl $R \gets \bigcup_{v \in V} D_v$ \label{line:rred} \;
}
\nl $\KwSty{return}~D$ \;
}
\caption{Variable initialisation for \cref{algorithm:sip}}
\label{algorithm:init}
\end{algorithm}

@@

\begin{algorithm}[h]
\DontPrintSemicolon
\SetKwSwitch{Switch}{Case}{Other}{case}{of}{}{otherwise}{endcase}{endsw}
\nl $\FuncSty{search}$ (GraphPairs $L$, Domains $D$) $\rightarrow$ Fail $F$ \KwSty{or} Success \;
\nl \Begin{
    \nl \lIf {$D = \emptyset$}{$\KwSty{return}~\textnormal{Success}$\label{line:success}}
    \nl $D_v \gets \textnormal{a domain in $D$ with minimum size, tiebreaking on static degree in}~\mathcal{P}$ \label{line:branch} \;
    \nl $F \gets \{ v \}$ \label{line:fcurrent} \;
    \nl \ForEach {$v' \in D_v~\textnormal{ordered by static degree in}~\mathcal{T}$\label{line:parforeach}}{
        \nl $D' \gets \FuncSty{clone}(D)$ \;
        \nl \Switch{$\FuncSty{assign}(L, D', v, v')$\label{line:assign}}{
            \nl \lCase{\textnormal{Fail $F'$~\KwSty{then}}}{$F \gets F \cup F'$\label{line:fassign}}
            \nl \Case{\textnormal{Success~\KwSty{then}}}{
                \nl \Switch{$\FuncSty{search}(L, D' - D_v)$\label{line:recursesearch}}{
                    \nl \lCase{\textnormal{Success~\KwSty{then}}}{$\KwSty{return}~\textnormal{Success}$}
                    \nl \Case{\textnormal{Fail $F'$~\KwSty{then}}\label{line:searchfailed}}{
                        \nl \lIf{$\nexists\,w \in F'~\KwSty{such that}~D_w \ne D'_w$}{$\KwSty{return}~\textnormal{Fail}~F'$\label{line:backjump}}
                        \nl $F \gets F \cup F'$ \label{line:fsub} \;
                    }
                }
            }
        }
    }
    \nl $\KwSty{return}~\textnormal{Fail}~F$ \;
}
\caption{Recursive search for \cref{algorithm:sip}}
\label{algorithm:search}
\end{algorithm}

@@

\begin{algorithm}[h]
\DontPrintSemicolon
\nl $\FuncSty{assign}$ (GraphPairs $L$, Domains $D$, Vertex $v$, Vertex $v'$) $\rightarrow$ Fail $F$ \KwSty{or} Success \;
\nl \Begin{
    \nl $D_v \gets \{ v' \}$ \label{line:assigngets} \;
\nl \ForEach{$D_w \in D - D_v$}{
    \nl $D_w \gets D_w - v'$ \label{line:removev} \;
    \nl \ForEach{$(P,\,T) \in L$}{
        \nl \lIf{$v \sim_P w$}{$D_w \gets D_w \cap \operatorname{N}(T, v')$\label{line:adj}}
    }
    \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\textnormal{Fail}~\{ w \}$\label{line:wipeout}}
}
\nl $\KwSty{return}~\FuncSty{countingAllDifferent}(D)$ \;
}
\caption{Variable assignment for \cref{algorithm:assign}}
\label{algorithm:assign}
\end{algorithm}

@@

\begin{algorithm}[h]
\DontPrintSemicolon
\nl $\FuncSty{countingAllDifferent}$ (Domains $D$) $\rightarrow$ Fail $F$ \KwSty{or} Success \;
\nl \Begin{
\nl $F \gets \emptyset$ \;
\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$ \;
\nl \ForEach{$D_v \in D$ \textnormal{from smallest cardinality to largest\label{line:eachdomain}}}{
    \nl $F \gets F + v$ \;
    \nl $D_v \gets D_v \setminus H$ \label{line:elimhall} \;
    \nl $(A,\,n) \gets (A \cup D_v,\,n + 1)$ \label{line:acc} \;
    \nl \lIf{$D_v = \emptyset~\KwSty{or}~|A| < n$}{$\KwSty{return}~\textnormal{Fail}~F$\label{line:failhall}}
    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\label{line:hall}}
}
\nl $\KwSty{return}~\textnormal{Success}$ \;
}
\caption{Counting-based all-different propagation}
\label{algorithm:cad}
\end{algorithm}

>>>>

\begin{abstract}
    The subgraph isomorphism problem involves finding a pattern graph inside a target graph.  We
    present a new bit- and thread-parallel constraint-based search algorithm for the problem, and
    experiment on a wide range of standard benchmark instances to demonstrate its effectiveness.  We
    introduce supplemental graphs, to create implied constraints. We use a new low-overhead, lazy
    variation of conflict directed backjumping which interacts safely with parallel search, and a
    counting-based all-different propagator which is better suited for large domains.
\end{abstract}

@@

This constraint-based search approach dates back to works by Ullmann \cite{Ullmann:1976} and
McGregor \cite{McGregor:1979}, and was improved upon in the LV \cite{Larrosa:2002} and VF2
\cite{Cordella:2004} algorithms.  More recently, ILF \cite{Zampelli:2010}, LAD \cite{Solnon:2010}
and SND \cite{Audemard:2014} are algorithms which take a ``deep thinking'' approach, using strong
inference at each stage of the search.  This is powerful, but we observe LAD or SND sometimes make
less than one recursive call per second with larger target graphs, and cannot always explore enough
of the search space to find a solution in time. This motivates an alternative approach: on the same
hardware, we will be making $10^4$ to $10^6$ recursive calls per core per second.  The main features
of our algorithm are:

@@

    \item We describe a clone-comparing variation of conflict-directed backjumping, which does not
        require conflict sets. We show that an all-different propagator can produce reduced
        conflict explanations, which can improve backjumping.

@@

\begin{proposition}\label{proposition:supplemental}
    Let $i : P \hookrightarrow T$ be a subgraph isomorphism. Then $i$ is also
    \begin{enumerate}[itemindent=1em,topsep=2pt,itemsep=0.2pt]
        \item a subgraph isomorphism $i^d : P^d \hookrightarrow T^d$ for any $d \ge 1$, and
        \item a subgraph isomorphism $i^{\left[c, l\right]} : P^{\left[c, l\right]} \hookrightarrow
            T^{\left[c, l\right]}$ for any $c, l \ge 1$.
    \end{enumerate}
\end{proposition}

@@

\begin{algorithm}[b]
\DontPrintSemicolon
\nl $\FuncSty{nonInducedSubgraphIsomorphism}$ (Graph $\mathcal{P}$, Graph $\mathcal{T}$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$|\operatorname{V}(\mathcal{P})| > |\operatorname{V}(\mathcal{T})|$}{$\KwSty{return}~\KwSty{false}$\label{line:enough}}
    \nl Discard isolated vertices in $\mathcal{P}$\label{line:isolated} \;
    \nl $L \gets \FuncSty{createSupplementalGraphList}(\mathcal{P}, \mathcal{T})$ \label{line:buildsup}\;
    \nl $D \gets \FuncSty{init}(\operatorname{V}(\mathcal{P}), \operatorname{V}(\mathcal{T}), L)$
    \label{line:initdomains}\;
    \nl \lIf{$\FuncSty{countingAllDifferent}(D) \ne \textnormal{Success}$}{$\KwSty{return}~\KwSty{false}$\label{line:initalldiff}}
    \nl $\KwSty{return}~\FuncSty{search}(L, D) = \textnormal{Success}$ \label{line:search} \;
}

@@

\begin{algorithm}[b]
\DontPrintSemicolon
\nl $\FuncSty{init}$ (Vertices $V$, Vertices $R$, GraphPairs $L$) $\rightarrow$ Domains \;
\nl \Begin{
    \nl \Repeat{\nl$R$ \textnormal{is unchanged}}{
        \nl \ForEach{$v \in V$}{
            \nl $D_v \gets \bigcap_{(P,\,T) \in L} \left\{ w \in R :~
                v\underset{P}{\sim}v \Rightarrow w\underset{T}{\sim}w \wedge \operatorname{S}(P, v) \preceq
            \operatorname{S}(T[R], w) \right\}$ \label{line:filter} \;
        }
    \nl $R \gets \bigcup_{v \in V} D_v$ \label{line:rred} \;
}
\nl $\KwSty{return}~D$ \;
}
\caption{Variable initialisation for \cref{algorithm:sip}}
\label{algorithm:init}
\end{algorithm}

@@

\begin{algorithm}[t]
\DontPrintSemicolon
\SetKwSwitch{Switch}{Case}{Other}{case}{of}{}{otherwise}{endcase}{endsw}
\nl $\FuncSty{search}$ (GraphPairs $L$, Domains $D$) $\rightarrow$ Fail $F$ \KwSty{or} Success \;
\nl \Begin{
    \nl \lIf {$D = \emptyset$}{$\KwSty{return}~\textnormal{Success}$\label{line:success}}
    \nl $D_v \gets \textnormal{a domain in $D$ with minimum size, tiebreaking on static degree in}~\mathcal{P}$ \label{line:branch} \;
    \nl $F \gets \{ v \}$ \label{line:fcurrent} \;
    \nl \ForEach {$v' \in D_v~\textnormal{ordered by static degree in}~\mathcal{T}$\label{line:parforeach}}{
        \nl $D' \gets \FuncSty{clone}(D)$ \;
        \nl \Switch{$\FuncSty{assign}(L, D', v, v')$\label{line:assign}}{
            \nl \lCase{\textnormal{Fail $F'$~\KwSty{then}}}{$F \gets F \cup F'$\label{line:fassign}}
            \nl \Case{\textnormal{Success~\KwSty{then}}}{
                \nl \Switch{$\FuncSty{search}(L, D' - D_v)$\label{line:recursesearch}}{
                    \nl \lCase{\textnormal{Success~\KwSty{then}}}{$\KwSty{return}~\textnormal{Success}$}
                    \nl \Case{\textnormal{Fail $F'$~\KwSty{then}}\label{line:searchfailed}}{
                        \nl \lIf{$\nexists\,w \in F'~\KwSty{such that}~D_w \ne D'_w$}{$\KwSty{return}~\textnormal{Fail}~F'$\label{line:backjump}}
                        \nl $F \gets F \cup F'$ \label{line:fsub} \;
                    }
                }
            }
        }
    }
    \nl $\KwSty{return}~\textnormal{Fail}~F$ \;
}
\caption{Recursive search for \cref{algorithm:sip}}
\label{algorithm:search}
\end{algorithm}

@@

\begin{algorithm}[t]
\DontPrintSemicolon
\nl $\FuncSty{assign}$ (GraphPairs $L$, Domains $D$, Vertex $v$, Vertex $v'$) $\rightarrow$ Fail $F$ \KwSty{or} Success \;
\nl \Begin{
    \nl $D_v \gets \{ v' \}$ \label{line:assigngets} \;
\nl \ForEach{$D_w \in D - D_v$}{
    \nl $D_w \gets D_w - v'$ \label{line:removev} \;
    \nl \ForEach{$(P,\,T) \in L$}{
        \nl \lIf{$v \sim_P w$}{$D_w \gets D_w \cap \operatorname{N}(T, v')$\label{line:adj}}
    }
    \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\textnormal{Fail}~\{ w \}$\label{line:wipeout}}
}
\nl $\KwSty{return}~\FuncSty{countingAllDifferent}(D)$ \;
}
\caption{Variable assignment for \cref{algorithm:assign}}
\label{algorithm:assign}
\end{algorithm}

@@

\begin{algorithm}[t]
\DontPrintSemicolon
\nl $\FuncSty{countingAllDifferent}$ (Domains $D$) $\rightarrow$ Fail $F$ \KwSty{or} Success \;
\nl \Begin{
\nl $(F,\,H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,\emptyset,\,0)$ \;
\nl \ForEach{$D_v \in D$ \textnormal{from smallest cardinality to largest\label{line:eachdomain}}}{
    \nl $F \gets F + v$ \;
    \nl $D_v \gets D_v \setminus H$ \label{line:elimhall} \;
    \nl $(A,\,n) \gets (A \cup D_v,\,n + 1)$ \label{line:acc} \;
    \nl \lIf{$D_v = \emptyset~\KwSty{or}~|A| < n$}{$\KwSty{return}~\textnormal{Fail}~F$\label{line:failhall}}
    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\label{line:hall}}
}
\nl $\KwSty{return}~\textnormal{Success}$ \;
}
\caption{Counting-based all-different propagation}
\label{algorithm:cad}
\end{algorithm}
